---
title: "Test API connection to FAIRDOMHub"
author: A. Blejec <andrej.blejec@nib.si> | National Institute of Biology | Ljubljana,
  Slovenia
date: "10 junij 2018"
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Available at [RPubs](http://rpubs.com/ablejec/396238)

# Introduction

I will try to connect to the FAIRDOM SEEK with the collection of their APIs. 

The API description with examples is available [here](https://app.swaggerhub.com/apis/FAIRDOM/SEEK/0.1#/people/updatePerson).

My final goal is to enable upload/download from pISA-tree to FAIRDOMHub. This feature will be a part of the R package `pISAR` that will enable use of pISA structure and metafile information for reproducible statistical analyses.

# Package `RCurl`

My first attempt will be the [RCurl](https://cran.r-project.org/package=RCurl) package avaialbe at CRAN. Additional [description](http://www.omegahat.net/RCurl/) is a part of Omegahat project.

```{r}
library(RCurl)
```

## Check existence

First check if the base URL exists

```{r}
(FAIRDOMExists = url.exists("https://www.fairdomhub.org/"))
```


## Get personal information

For this I will use the `GET \people\{id}` from documentation and my ID {808}.

```{r}
person <- getURL("https://www.fairdomhub.org/people/808")
str(person)
```

It returns the HTML content and might not be the optimal way.

# Package `httr`

The use is described in [LockeData blog](https://itsalocke.com/blog/r-quick-tip-microsoft-cognitive-services-text-analytics-api/).

This solution uses packages `httr` and `jsonlite` and is based on `libcurl` (the same is used by RCurl). Package `httr` has useful vignettes that were a good guide for the example below.

```{r}
library(httr)
library(jsonlite)
```

## Get personal information

```{r}
r <- GET("https://www.fairdomhub.org/people/808",
         add_headers(Accept="application/json"))
names(r)
r
```
The response is structured and can be investigated with the helper functions. The response has three important parts: status, headers, and body.

### Status

Check status

```{r}
http_status(r)
```
or just a status code

```{r}
r$status_code
```

### Body

Body can be accessed with the function `content`

```{r}
cont <- content(r,"text")
cont
length(cont)
```

This gives us a text version of the JSON object. We can use built-in parsers to get other forms of data. We can parse the content into the R list object

```{r}
cont <- content(r,"parsed",type="application/json")
cont
```
### The structure of the parsed response

```{r}
names(cont)
cont$jsonapi
names(cont$data)
```

Response elements

```{r}
cont$data$id
```

List the response content

```{r}
d <- cont$data
for (nm in names(d)) {
  cat("\n",nm,"--------------\n\n")
  print(d[[nm]])
}
```

Person data are in the *attributes* part

Here is the first name of the person with id 808:

```{r}
cont$data$attributes$first_name
```

# Conclusion

The rudimetary R read connection with FAIRDOMHub can be established using the package `httr`. From here it should be fairly possible to write the wrapper functions for easier usage.

## Wrapper functions for `GET` and `content`

```{r}
# this function is equivaalent to a call:
# curl -X GET "https://www.fairdomhub.org/what/id" -H  "accept: application/json"
#
myGET <- function(what, id,
                  uri="https://www.fairdomhub.org", ... ){
  if(!missing(what)) uri <- paste0(uri,"/",what)
  if(!missing(id)) uri <- paste0(uri,"/",id)
  r <- GET(uri,
         add_headers(Accept="application/json"))
  cat("Status code:",r$status_code,"\n")
  invisible(r)
}

myData <- function(r,
                  part="attributes",
                  type="application/json",  ...){
  invisible(content(r,"parsed",type=type)$data[[part]])
}
myDatas <- function(r,part,
                  type="application/json",  ...){
  if(missing(part)) invisible(content(r,"parsed",type=type)$data) else
  invisible(content(r,"parsed",type=type)$data[part])
}

```

## Use of wrapper functions

Get data for person 808

```{r}
# https://www.fairdomhub.org/people/808
r <- myGET("people","808")
r$status_code
```

Get the person 808 last name and tools

```{r}
myData(r)$last_name
myData(r)$tools
```

Get person 808 institution id

```{r}
myData(r,"relationships")$institutions
instID <- myData(r, "relationships")$institutions$data[[1]]$id
instID
```

Get institution \r instID` name. The details are in the *institutions* part.

```{r}
# https://www.fairdomhub.org/institutions/189
rinst <- myGET("institutions",instID)
myData(rinst)$title
```

Full institution description

```{r}
catln <- function(...) cat(...,"\n")
d <- sapply(myData(rinst),catln)
```


### Search for person

```{r}
r <- myGET("people")
d <- myDatas(r)
length(d)
names(d[[1]])
titles <- sapply(d,function(x) x$attributes$title)
head(titles)
d[[which(titles=="Andrej Blejec")]]
id <- d[[which(titles=="Andrej Blejec")]]$id
id
```




